Secure Composition of Robust and Optimising Compilers

### 2023-07-10 popl/main.tex
sec 6

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, there is only a passive attacker model that observes the execution and, e.g., retreives secrets from that.
To consider a stronger model similarily to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the concept of satisfaction is extended with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., programs with free variables, which need to be defined prior to execution.
Programs with free variables can get stuck during execution.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in not only secure compilation~\cite{abate2019jour,devriese2018parametricity}, %
% \MP{others do this too, parametricity results e.g.}
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
%\MPin{
% the behaviour of the result?
%%%%%%%    MK: def:propsat defines satisfaction in terms of programs, no need to talk about behavior here
%}

@@

\begin{center}
  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
  \typerule{ms-step}{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
  }{
    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
  }{ms-step}
\end{center}
The combined monitor runs the one for \gls{tms} and the one for \gls{sms} in a lockstep. The trace agreement similarily just relates a specification event with an abstract \gls{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls{sms}-event $\abstrace[_{\smssafe}]$.
% \begin{center}
%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{ms}.''}$\;$\\
%   %
%   \typerule{ms-authentic}{
%     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
%     \smstraceagree[]{\event}{\absevent_{\smssafe}}
%   }{
%     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
%   }{ms-authentic}
% \end{center}

@@

      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
      \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
    \end{tikzpicture}
  \end{center}
  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}.}\label{fig:pipeline}
\end{figure}

This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
\MPin{
	explain how to read the figure: arrows mean, annotations mean ... ...
}
It\MP{what} demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
Since these properties compose into \MP{}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
The last step of the chain ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
The final result is that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).

\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
\MPin{
	this environment should just be a align*
}
  $$
  \begin{array}{rcl}
%    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) & = & \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
                                                                                                \left[\cca(\src{\expr})\right] %
                                                                                                 }{\labort[\trg]}}
  \end{array}
  $$

@@

\MPin{
	this should be a subsection, not a proof sketch.
}

\begin{proof}[Proof Sketch]
The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
\MPin{
	explain how to read the figure
}
Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace-based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
\MPin{
	what does this mean?
	what is a backtranslation?
}
\MPin{
	why can we use a BT?
	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
	all these intuitions are missing.
	they are technical, so we only keep them here.
}
\Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.
Well-typing\MP{well-typedness} means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$-level execution with the $\src{L_{\tmssafe}}$-level execution.

For the $\trg{L}$-execution, the compiler wrapper\MP{undefined terminology. call it like this before} code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
In this specific case, the compiler needs not insert additional code after the component is done with its execution, but a symmetric situation can happen here.\MP{vague and imprecise}
Similarly, the backtranslation may also insert wrapper code prior and after switching to and from the component to the context. \MP{wut? why is a wrapper needed?}
\MKin{
  more detail or not? ;
  i.e., explain $\multimap$ and $\approx$?
  \MPin{
  yes
  }
}
\MPin{
	the general idea that later proofs use the same structure is also not here.
}
\end{proof}

@@

Up to getting and setting memory\MP{reading and writing? uniform lingo}, the compiler\MP{notation} is the identity function, i.e., all omitted cases are a recoloring of $\trg{L}$ to $\Lms$.\MP{reword as before}
For memory accesses, however, the compiler inserts a bounds-check that enforces \gls{sms}.
To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
Again, linking this back to the \texttt{strncpy} example corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
\MPin{
	example code?
}

@@

It is now possible to use \Cref{thm:rtp} to extract a compiler from $\Ltms$ to $\Lms$ that is \gls{ms} by applying \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms}.
\MP{
	bad sentence. rewrite.
}

@@

\begin{center}
  $$
  \begin{array}{rcll}
    \cccf(\irl{\llet{x}{n}{\expr}}) & = & \cccf(\irl{\expr\subst{n}{x}}) &\\
    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = & \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}\\
  \end{array}
  $$
\end{center}
\MPin{
	substitution are a runtime semantics thing.
	what you want is a compiler with 2 arguments, where in one you keep track of the static substitions to perform,
	and 2 cases on how to handle variables: if they're in the second arg, then substitute, otherwise don't
}

The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls{dce} and \gls{cf}, respectively.
Both are secure with respect to \gls{ms}, the proof of which is relatively easy, since the backtranslation can use the original context without modification and both transformations do not change anything with regards to allocation, use, or deallocation.
\MPin{
	the point is different: the proof is easy because the transofrmations do not change the way memory is touched,
	and from a security perspective, the attacker has no extra power since the target is the same as in the source.
	.
	rewrite.
}
That is, the order of appearance of the events on the execution trace stays the same.

\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls{ms}]\label{thm:ccdce:rtp:ms}
  $\rtp{\ccdce}{\mssafe}$ % \Coqed
\end{theorem}
\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls{ms}]\label{thm:cccf:rtp:ms}
  $\rtp{\cccf}{\mssafe}$ % \Coqed
\end{theorem}

@@

Given the fact that $\Lscct$ provides a \gls{cct}-mode, the compiler has to ensure that execution in the component always happen in this mode.
\MPin{
	reuse the word 'wrapper' here
}
The context can overwrite the flag and exit the mode, but upon invoking a callable object\MP{again with the word object} that is part of the component, the flag would be set again.
Because of this, the compiler is secure with respect to \gls{scct}, similarily\MP{misspelled} proven as in \Cref{subsec:cs:tms}.

@@

Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem can be shown\MP{holds}.

\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ % \Coqed
\end{theorem}
\MPin{
This you can write the proof of.
Probably even in coq
}

@@

This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarily to how parser combinators work.
%To this end, it may be necessary to extend existing frameworks for multi-language semantics.
It is also interesting to extend our case-study and verify all results of it in Coq.
This way, it is possible to obtain an executable, secure compiler.
However, the formalisation effort for backtranslation proofs is known to be enormous, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.

>>>>

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, there is only a passive attacker model that observes the execution and, e.g., retreives secrets from that.
To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the concept of satisfaction is extended with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., programs with free variables, which need to be defined prior to execution.
Programs with free variables can get stuck during execution.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in not only secure compilation~\cite{abate2019jour,devriese2018parametricity}, %
% \MP{others do this too, parametricity results e.g.}
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the result of which must satisfy $\pi$}.
%\MPin{
% the behaviour of the result?
%%%%%%%    MK: def:propsat defines satisfaction in terms of programs, no need to talk about behavior here
%}

@@

\begin{center}
  \judgbox{\monitorcheck[]{\msmonitor}{\msmonitor[']}{\absevent}}{,,Monitor $\msmonitor$ does one step to $\msmonitor[']$ given event $\absevent$.''}$\;$\\
  \typerule{ms-step}{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent_{\tmssafe}} &
    \monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent_{\smssafe}} &
  }{
    \monitorcheck[]{(\tmsmonitor,\smsmonitor)}{(\tmsmonitor['],\smsmonitor['])}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
  }{ms-step}
\end{center}
The combined monitor runs the one for \gls{tms} and the one for \gls{sms} in a lockstep. The trace agreement similarly just relates a specification event with an abstract \gls{tms}-event $\abstrace[_{\tmssafe}]$ and with an abstract \gls{sms}-event $\abstrace[_{\smssafe}]$.
% \begin{center}
%   \judgbox{\mstraceagree{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{ms}.''}$\;$\\
%   %
%   \typerule{ms-authentic}{
%     \tmstraceagree[]{\event}{\absevent_{\tmssafe}} &
%     \smstraceagree[]{\event}{\absevent_{\smssafe}}
%   }{
%     \mstraceagree[]{\event}{(\absevent_{\tmssafe},\absevent_{\smssafe})}
%   }{ms-authentic}
% \end{center}

@@

      \draw[dashed] ($(sectms)-(0.8,0)$) -- ($(sectms)-(0.8,1.25)$);
      \draw[dashed] ($(sectms)-(-0.9,0)$) -- ($(sectms)-(-0.9,1.25)$);
      \draw[dashed] ($(secsms)-(-1.2,0)$) -- ($(secsms)-(-1.2,1.25)$);
      \draw[dashed] ($(secscct)-(1.2,0)$) -- ($(secscct)-(1.2,1.25)$);
      \draw[dashed] ($(secscct)-(-1.1,0)$) -- ($(secscct)-(-1.1,1.25)$);
    \end{tikzpicture}
  \end{center}
  \caption{Visualisation of the optimising compilation pipeline that attains a combination of \gls{ms} and \gls{cct}. %
    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
    All edges are secure compilers, but dotted edges use the presented framework (\Cref{sec:compcomp}) and strikethrough edges classic proof techniques. %
    The dashed lines partition the graph into the sections where the respective theorems are presented.
  }\label{fig:pipeline}
\end{figure}

This section defines several secure compilers, each of which robustly preserves a different property of interest as depicted in \Cref{fig:pipeline}.
The section demonstrates the power of the framework (\Cref{sec:compprop,sec:compcomp}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls{msscct}.
The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls{tms} (\Cref{thm:cca:rtp:tms}).
From here, an instrumentation from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls{sms} (\Cref{thm:ccb:rtp:sms}).
Since these properties compose into \gls{ms}, composing these passes yields a compiler that robustly preserves \gls{ms} (\Cref{thm:ccab:rtp:ms}).
At this stage, the section presents two optimising translations, namely \gls{cf} and \gls{dce}, each of which robustly preserves \gls{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
These translations can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
The last step of the chain ensures that code stays \gls{scct} (\Cref{thm:ccscct:rtp:scct}) when lowered from $\Lms$ to $\Lscct$.
The final result is that the whole compilation chain robustly preserves \gls{msscct} (\Cref{thm:ccall:rtp:msscct}).

\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
  \begin{align*}
%    \cca(\src{\lbinop{\expr[_{1}]}{\expr[_{2}]}}) & = & \lbinop[\trg]{\left[\cca(\src{\expr[_{1}]})\right]}{\left[\cca(\src{\expr[_{2}]})\right]} \\ [0.33cm]
    \cca(\src{\lfunction{g}{x:\natt\to\type[_{e}]}{\expr}}) = \lfunction[\trg]{\trg{g}}{\trg{x}}{\lifz[\trg]{\trg{\lhast{x}{\natt}}}{%
                                                                                                \left[\cca(\src{\expr})\right] %
                                                                                                 }{\labort[\trg]}}
  \end{align*}

@@

\subsubsection{Proof Sketch of Robust Temporal Memory Safety Preservation}

The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
\MPin{
	explain how to read the figure
}
Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace-based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
\MPin{
	what does this mean?
	what is a backtranslation?
}
\MPin{
	why can we use a BT?
	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
	all these intuitions are missing.
	they are technical, so we only keep them here.
}
\Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.
Well-typing\MP{well-typedness} means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$-level execution with the $\src{L_{\tmssafe}}$-level execution.

For the $\trg{L}$-execution, the compiler wrapper\MP{undefined terminology. call it like this before} code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
In this specific case, the compiler needs not insert additional code after the component is done with its execution, but a symmetric situation can happen here.\MP{vague and imprecise}
Similarly, the backtranslation may also insert wrapper code prior and after switching to and from the component to the context. \MP{wut? why is a wrapper needed?}
\MKin{
  more detail or not? ;
  i.e., explain $\multimap$ and $\approx$?
  \MPin{
  yes
  }
}
\MPin{
	the general idea that later proofs use the same structure is also not here.
}

@@

The compiler $\cc{\trg{L}}{\Lms}$ only inserts bounds-checks to enforce \gls{sms} whenever reading from or writing to memory.
To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.
Again, linking this back to the \texttt{strncpy} example (\Cref{ex:strncpy:sms}) corresponds to the insertion of bounds check right in front of the accesses to \texttt{y[i]} and \texttt{x[i]}.
In turn, the event $\ev{Use}\ \loc_{x}\ 2;\comp;\unlock$ would not be emitted during execution.
\MPin{
	example code?
    MK: still needed now that we reference ex. 3.10?
}

@@

\Cref{thm:ccab:rtp:ms} follows from \Cref{thm:cca:rtp:tms,thm:ccb:rtp:sms} using \Cref{thm:rtp}.

@@

\begin{center}
  $$
  \begin{array}{rcll}
    \cccf(\irl{\llet{x}{n}{\expr}}) & = & \cccf(\irl{\expr\subst{n}{x}}) &\\
    \cccf(\irl{\lbinop{n_{1}}{n_{2}}}) & = & \irl{n_{3}} & \text{where }\irl{n_{3}}\text{ is }\lbinop{\irl{n_{1}}}{\irl{n_{2}}}\\
  \end{array}
  $$
\end{center}
\MPin{
	substitution are a runtime semantics thing.
	what you want is a compiler with 2 arguments, where in one you keep track of the static substitions to perform,
	and 2 cases on how to handle variables: if they're in the second arg, then substitute, otherwise don't
  %
  MK: No? I get how they are used as part of the runtime semantics, but a compiler can just as well make use of a substitution function that models the runtime semantics. I see how these two are different. However, I think it's easier in this case to be a bit inprecise, since adding another argument to the compiler breaks our compositionality framework (unless it's hidden with a monad...) so, IMO, we should view substitution here as a meta-level, but still language-dependent, function that the compiler has access to. If we don't do this, we essentially re-encode the definition of subst into the compiler
}

The two optimising compiler passes from $\Lms$ to $\Lms$ perform \gls{dce} and \gls{cf}, respectively.
Both are secure with respect to \gls{ms}, the proof of which is relatively easy, because \gls{dce} and \gls{cf} do not change the way memory accesses happen.
Moreover, since the input and output languages to these compilers are the same, attacker contexts do not have more power in the target language than in the source.

\begin{theorem}[Compiler $\ccdce$ is secure with respect to \gls{ms}]\label{thm:ccdce:rtp:ms}
  $\rtp{\ccdce}{\mssafe}$ \Coqed
\end{theorem}
\begin{theorem}[Compiler $\cccf$ is secure with respect to \gls{ms}]\label{thm:cccf:rtp:ms}
  $\rtp{\cccf}{\mssafe}$ \Coqed
\end{theorem}

@@

Given the fact that $\Lscct$ provides a \gls{cct}-mode that can be turned on or off, the compiler inserts wrapper code for function bodies to ensure that execution in the component always happen in this \gls{cct}-mode.
The context can overwrite the flag and exit the mode, but upon invoking a function that is part of the component, the flag would be set again.
Because of this, the compiler is secure with respect to \gls{scct}, similarly proven as in \Cref{subsec:cs:tms}.

@@

Let $\ccmsscct$ be the compiler that is the composition of $\cca$, $\ccb$, $\cccf$, $\ccdce$, and $\ccscct$, then the following theorem holds.

\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
\end{theorem}
\MPin{
This you can write the proof of.
Probably even in coq
%
MK: Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}
}

@@

This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
%To this end, it may be necessary to extend existing frameworks for multi-language semantics.
It is also interesting to extend our case-study and verify all results of it in Coq.
This way, it is possible to obtain an executable, secure compiler.
However, the formalisation effort for backtranslation proofs is known to be enormous, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
